---
title: "HW 4, pt. 1: Programming Exercises"
author: "Julia Du"
output: github_document
---

## Load necessary libraries

```{r}
library(tidyverse)
library(knitr)
library(ggthemes)

theme_set(theme_minimal())

#not relevant - class exercises
mean_mtcars <- vector("numeric", length = ncol(mtcars))
for(i in seq_along(mtcars)) {
  output[[i]] <- mean(mtcars[[i]])
}

max_diamonds <- vector("numeric", length = ncol(diamonds))
for(i in seq_along(diamonds)) {
  output[[i]] <- max(diamonds[[i]])
}
```

## Compute the number of unique values in each column of `iris`

### Using a `for` loop

```{r}
as_tibble(iris)

unique_iris <- vector("numeric", length = ncol(iris)) 
names(unique_iris) <- names(iris)
for(i in seq_along(iris)) {
  unique_iris[[i]] <- length(unique(iris[[i]]))
}
#alternative body is: n_distinct(iris[[i]])

unique_iris
```

### Using a `map` function

```{r}
iris %>%
  map_int(n_distinct)
#alternative is map_int(iris, function(x) length(unique(x)))
```

## Calculate the square of each element in vector `x`

```{r}
x <- seq(from = 30, to = 1)
x
```

### Using a `for` loop

```{r}
squaredx <- vector("numeric", length = length(x)) 
for(i in seq_along(x)) {    #could also use for(i in 1:length(x))
  squaredx[[i]] <- (x[[i]]) ^2
}
squaredx
```

### Using a `map` function

```{r}
x %>%
map_dbl(~(.x)^2) 
#alternative is map_dbl(x, function(x) x^2)
#my tilde in the 2nd line takes the place of function(x) in 3rd line
```

## Write a function to calculate length of sides in a right-triangle using the Pythagorean Theorem

```{r}
pyth1 <- function(a = NULL, b = NULL, c = NULL) {
  sides <- sort(c(a, b, c)) #so sorting smallest to largest
  if(!is.numeric(sides)) {
    stop("sides must be numeric") 
  } else if(length(sides) != 2) {
    stop("provide only 2 values")
  } else if (is.null(a)) {              
    sqrt(sides[2]**2 - sides[1]**2)   #satisfies pyth1( , 5, 4)
  } else if (is.null(b)) {
    sqrt(sides[2]^2 - sides[1]^2)       
  } else if (is.null(c)) {
    sqrt(a**2 + b**2)
  }
} 

#test examples
pyth1(a = 3, b = 4, ) 
pyth1(3, , 5)
pyth1(5, , 4)
pyth1(, 5, 4)
pyth1(, 4, 5)

#gives planned errors
pyth1(3,4,5)
pyth1("one", 2)
```

If pyth1's inputs are not a AND b (wherein pyth1 will proceed to calculate the hypotenuse), then pyth1 knows that one of the sides it's been given is already the hypotenuse (the longest side).

## Session info

```{r}
devtools::session_info()
```


