---
title: "HW 4, pt. 2: Exploring World Bank Data"
author: "Julia Du"
output: github_document
---

## Load necessary libraries

```{r}
library(tidyverse)
library(ggthemes)
library(ggridges)
library(viridis)
library(forcats)
library(readr)
library(knitr)

theme_set(theme_minimal())
```

## Write a function to import the data files

  I'm interested in the following four variables: 
    
    1. Total population growth (annual %) = SP.POP.GROW
    2. Urban population growth (annual %) = SP.URB.GROW
    3. Population density (people per sq. km of land area) = EN.POP.DNST
    4. Life expectancy at birth, total (years) = SP.DYN.LE00.IN
    
I chose these indicators because I think they could influence each other.

```{r function_chunk}
import_tidydata <- function(file_name) {
 path = "./data_world_bank/"  
 read_csv(paste0(path, file_name), skip = 4) %>%
   rename(country_name = `Country Name`, country_code = `Country Code`, indicator_name = `Indicator Name`, indicator_code = `Indicator Code`) %>%
   filter(indicator_code %in% c("SP.POP.GROW", "SP.URB.GROW", "EN.POP.DNST", "SP.DYN.LE00.IN")) %>%
   #selecting chosen 4 variables: population growth, urban population growth, population density, & life expectancy at birth. doing it here & not in final select of function so that function (theoretically) runs faster
   select(-c(country_code, indicator_name)) %>%
   pivot_longer(cols = `1960`:`2019`, names_to = "year", values_to = "values") %>%
   pivot_wider(names_from = indicator_code, values_from = values) %>%
   select(-X65)
}

#This is what I used to check my work while creating the function.
sampledfname <- "API_ABW_DS2_en_csv_v2_1226885.csv"
sampledf <- import_tidydata(sampledfname)
```

## Import the data

```{r import, message = FALSE, cache = TRUE, dependson = "function_chunk"}
# used message = FALSE to avoid pages of same message saying it created a new column [X65], as seen above
masterdata <- list.files(path = "./data_world_bank/", pattern = "\\.csv$") %>% #could also write "*.csv"
map(import_tidydata) %>%
  reduce(rbind) %>%
    mutate(country_name = if_else(country_name == "Korea, Rep.", "South Korea", country_name)) 
#changing name of South Korea now to set up for later analysis
```

## Explore the data
# **East Asian Countries** 
My initial analysis focuses on 3 East Asian countries: China, Japan, and South Korea.
```{r}
focus_data <- masterdata %>%
  filter(country_name %in% c("China", "Japan", "South Korea")) %>%
  rename(totalpopgrow = `SP.POP.GROW`, urbpopgrow = `SP.URB.GROW`, popdensity = `EN.POP.DNST`, life_expectancy = `SP.DYN.LE00.IN`)

focus_data %>%
  group_by(country_name) %>%
  drop_na(totalpopgrow:life_expectancy) %>%
  summarize(
    mean(totalpopgrow), mean(urbpopgrow), mean(popdensity), mean(life_expectancy)
    ) %>%
  knitr::kable(caption = "Averages of indicators across East Asian countries, from 1960-2020", col.names = c("Country", "Total population growth (%)", "Urban population growth (%)", "Population density (per sq. km)", "Life expectancy (yrs)"), digits = 2)
```

For all 3 selected East Asian countries, the table shows the average values of the indicators of interest from 1960 to 2020. On average, China experienced more population growth – both total and urban – over the years. However, South Korea has the biggest average population density, nearly 4x China’s average over the years. Japan had the highest average life expectancy at birth, roughly 78 years, which is 10 years more than China’s average life expectancy.


# **3 Geographical Regions (9 total countries)** 
I started my analysis just focusing on the 3 East Asian countries above, but I realized I could look at *regions*. I focused on 3 geographical regions: East Asia (same as above), North America (Mexico, Canada, US), and Western Europe (France, Germany, UK). 
```{r}
#compare geographical regions
require(forcats)

e_asia <- c("China", "Japan", "South Korea")
n_ame <- c("Mexico", "Canada", "United States")
w_eur <- c("France", "Germany", "United Kingdom")

region_level <- c(e_asia, n_ame, w_eur)
  #"East Asia", "North America", "Western Europe")

regional_data <- masterdata %>%
  filter(country_name %in% c(e_asia, n_ame, w_eur)) %>%
  rename(totalpopgrow = `SP.POP.GROW`, urbpopgrow = `SP.URB.GROW`, popdensity = `EN.POP.DNST`, life_expectancy = `SP.DYN.LE00.IN`) %>%
  mutate(region = case_when(
    country_name %in% e_asia ~ "East Asia",
    country_name %in% n_ame ~ "North America",
    country_name %in% w_eur ~ "Western Europe"
  )) %>%
  mutate(country_name = factor(country_name, levels = region_level), country_name = fct_rev(country_name))

#table of mean indicators across region
regional_data %>%
  group_by(region) %>%
  drop_na(totalpopgrow:life_expectancy) %>%
  summarize(
    mean(totalpopgrow), mean(urbpopgrow), mean(popdensity), mean(life_expectancy)
    ) %>%
  knitr::kable(
    caption = "Averages of indicators across region, from 1960-2020", 
    col.names = c("Region", "Total population growth (%)", "Urban population growth (%)", "Population density (people per sq. km)", "Life expectancy (yrs)"), 
    digits = 2)

# violinplot of life_expectancy - all countries
regional_data %>%
  ggplot(mapping = aes(x = country_name, y = life_expectancy, fill = region)) +
  geom_violin() +
  geom_boxplot(width=0.1, color="grey", alpha=0.2) +
    labs(title = "Life expectancy in selected countries, 1960-2020", x = "", y = "Life expectancy at birth (yrs)", caption = "Source: The World Bank", fill = "Region") +
  theme(legend.position = "bottom") +
  coord_flip()

#violin plot of regional life expectancy
ggplot(data = regional_data, mapping = aes(x = region, y = life_expectancy, fill = region)) +
  geom_violin() +
  geom_boxplot(width=0.1, color="grey", alpha=0.2) +
    labs(title = "Annual life expectancy in selected regions, 1960-2020", x = "", y = "Life expectancy at birth (yrs)", caption = "Source: The World Bank", fill = "Country") +
  scale_y_continuous(limits = c(40, 90)) +
  theme(legend.position = "none")

#ridged population density - all countries
ggplot(data = regional_data, mapping = aes(x = popdensity, y = country_name, fill = region)) +
  geom_density_ridges(alpha = 0.6) +
  scale_fill_viridis(discrete = TRUE) +
  scale_color_viridis(discrete = TRUE) +
    labs(title = "Population density in selected countries, 1960-2020", x = "Population density (people per sq. km)", y = "", caption = "Source: The World Bank", fill = "Country") +
  theme(legend.position = "none")

#ridged pop density - 
ggplot(data = regional_data, mapping = aes(x = popdensity, y = fct_rev(region), fill = region)) +
  geom_density_ridges(alpha = 0.6) +
  scale_fill_viridis(discrete = TRUE) +
  scale_color_viridis(discrete = TRUE) +
    labs(title = "Population density in selected regions, 1960-2020", x = "Population density (people per sq. km)", y = "", caption = "Source: The World Bank", fill = "Country") +
  theme(legend.position = "none")
```

I first created a similar table to the East Asian table for these 3 geographical regions. The East Asian countries had a much higher average population density and a higher average annual urban population growth rate over 1960-2020. I didn’t expect this, but North America had the highest average annual total population growth rate. Finally, Western Europe had the highest average life expectancy at roughly 76 years. 
	
*I then created two violin plots to analyze life expectancy from our data covering 1960-2020.* The **first violin plot** looks at all 9 countries. Upon first glance, China is skewed a lot to the left and has the lowest median life expectancy (a little under 70 years). This seems logical, as China was largely rural with limited medical services prior to the Communist Revolution and is still undergoing rural revitalization programs today. Mexico is also skewed to the left, and the other East Asian countries seem to also be slightly left-skewed. Out of all the countries, Japan does seem to have the highest median life expectancy over the time range this data was collected. Interestingly, the Western European countries seems to be relatively centered and not skewed. The Western European countries, alongside the US & Canada, seem to have pretty similar median life expectancies. 
The **second violin plot** consolidates some of the trends described above: East Asian countries are left-skewed, while Western Europe is more centered. North America is also skewed to the left, likely influenced largely by Mexico’s left-skewness. Overall, Western Europe has the highest regional median life expectancy, and East Asia has the lowest. 

*I also created 2 ridgeline plots to analyze population density.* 
The **1st ridgeline plot** covers all 9 countries. We see a clustering of North American countries at low population densities, roughly 50 people per sq. km. The Western European countries have comparatively higher densities and are less clustered together, with France having noticeably lower population densities than its peers. The East Asian countries are even less clustered together, with China surprisingly centered around a population density of 100-125 people per sq. km. South Korea has a lot of spread and lacks a clear center population density over the years.
The **2nd ridgeline plot** consolidates the above trends. Since the North American countries were very close together, there is a clear center population density (which is the lowest out of all 3 regions). Western Europe has 2 peaks, since France had a much lower population density than the UK & Germany. East Asia has many peaks: all 3 countries had multiple peaks and South Korea in particular was extremely spread out, so there is no clear peak. East Asia does have some of the highest population densities, but it also had some very low densities, even overlapping a little into North America’s low densities. 


# **Indicators over Time - 3 geographical regions**
``` {r}
#line graphs of 3 regions, tracking variables over time

#population density over time
regional_data %>%
  ggplot(mapping = aes(x = year, y = popdensity, color = fct_rev(country_name), group = country_name)) +
  geom_line() +
  facet_wrap(region ~ .)  +
  labs(title = "Annual population density, 1960-2020", subtitle = "In selected countries", x = "Year", y = "Population density (people per sq. km)", caption = "Source: The World Bank", color = "Country") +
  scale_x_discrete(breaks = seq(1960, 2020, by = 10)) +
  theme(legend.position = "bottom")

#life expectancy over time
regional_data %>%
  ggplot(mapping = aes(x = year, y = life_expectancy, color = fct_rev(country_name), group = country_name)) +
  geom_line() +
  facet_wrap(region ~ .)  +
  labs(title = "Life expectancy, 1960-2020", subtitle = "In selected countries", x = "Year", y = "Life expectancy, at birth (yrs)", caption = "Source: The World Bank", color = "Country") +
  scale_x_discrete(breaks = seq(1960, 2020, by = 10)) +
  theme(legend.position = "bottom")

#total population growth rates over time
regional_data %>%
  ggplot(mapping = aes(x = year, y = totalpopgrow/100, color = fct_rev(country_name), group = country_name)) +
  geom_line() +
  facet_wrap(region ~ .)  +
  labs(title = "Annual total population growth, 1960-2020", subtitle = "In selected countries", x = "Year", y = "Total population growth", caption = "Source: The World Bank", color = "Country") +
  scale_x_discrete(breaks = seq(1960, 2020, by = 10)) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "bottom")

#comparing total pop growth to urban pop grow over time
ggplot(data = regional_data, mapping = aes(x = year)) +
  geom_line(mapping = aes(y = totalpopgrow/100, color = "Total population growth"), group = 1) +
  geom_line(mapping = aes(y = urbpopgrow/100, color = "Urban population growth"), group = 1) +
  facet_wrap(fct_rev(country_name) ~.) +
    labs(title = "Total & Urban Population Growth, 1960-2020", x = "Year", y = "Percent of annual population growth", caption = "Source: The World Bank", color = "Legend") +
  scale_x_discrete(breaks = seq(1960, 2020, by = 10)) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "bottom")
```

_I continued to analyze my indicators of interest, this time using **line graphs to see how they evolved over time.**_

Most of the 9 countries seemed to at least somewhat increase their **population densities** over time, though Canada barely changed its extremely low density over time. Germany also stays relatively stagnant, hovering around a density of 225 people per sq. km. South Korea in particular has a giant increase in population density, going from 250 people per sq. km in 1960 to 525 people per sq. km about sixty years later. East Asian countries (namely Japan and South Korea) clearly have higher population densities, while North American countries have the lowest (as noted in the earlier ridgeline graphs).
	
We can see that Western Europe's **life expectancies** have roughly stayed together over time, whereas North American countries were slightly diverged (though they are getting closer as time goes on). The East Asian countries’ life expectancies are even more divergent. In all regions here, we see an increase in life expectancy over time, with East Asian countries experiencing the most dramatic change (as China and South Korea started with noticeably lower expectancies). Today, we see roughly similar life expectancies for all countries, hovering around 80 yrs. Japan and South Korea (with modern life expectancies around 82 years) have actually surpassed the North American and Western European countries, while China and Mexico lag behind (around 75 years).
For all countries, we see a general slowing in the **total population growth rate** over time, but the rate stays positive for the most part. However, China seems to have decreased its population in the early 1960s (i.e. a negative growth rate), while Germany also experienced decreases in the early 2010s (due to a declining birth rate). Today, Japan, South Korea, France, and Germany – which have all been in the news for their declining birth rates – have very low total population growth rates, roughly less than 0.5%. Mexico has the highest total population growth rate today, a little over 1%, but that is still much lower than its historical rates. Historically, Western Europe seemed to have lower rates than the other 2 regions, but East Asia’s growth rates have decreased so much over the years that in modern times, East Asia and Western Europe have similar total population growth rates.
In general, it seems that **urban and total population growth rates** have both been slowing over time, but urban growth rates typically stay a bit higher than total growth rates. The total and urban growth rates of Germany, the UK, and the US track quite closely together. South Korea had a noticeably higher urban rate than its total rate from 1960 to 1990. Strangely, China’s urban rate dipped below its total population rate from 1965 to 1975 – this may be due to Cultural Revolution programs, one of which compelled (urban-based) scholars to work in the countryside. 



**Indicator Relationships**
```{r}
#relationships between indicators
ggplot(data = regional_data, mapping = aes(x = totalpopgrow/100, y = urbpopgrow/100)) +
  geom_line() +
  facet_wrap(fct_rev(country_name) ~.) +
    labs(title = "Total vs. Urban Population Growth, 1960-2020", x = "Percent of total population growth", y = "Percent of urban population growth", caption = "Source: The World Bank", color = "Country") +
  scale_y_continuous(labels = scales::percent) + 
  scale_x_continuous(labels = scales::percent) 

ggplot(data = regional_data, mapping = aes(x = life_expectancy, y = totalpopgrow/100)) +
  geom_line() +
  facet_wrap(fct_rev(country_name) ~.) +
    labs(title = "Life Expectancy vs. Total Population Growth, 1960-2020", x = "Life expectancy (yrs)", y = "Percent of total population growth", caption = "Source: The World Bank", color = "Country") +
  scale_y_continuous(labels = scales::percent) 

ggplot(data = regional_data, mapping = aes(x = life_expectancy, y = popdensity, color = fct_rev(country_name), group = country_name)) +
  geom_line() +
#  facet_wrap(fct_rev(country_name) ~.) +
    labs(title = "Life Expectancy vs. Population Density, 1960-2020", x = "Life expectancy (yrs)", y = "Population density (people per sq. km)", caption = "Source: The World Bank", color = "Country")
```

As expected and noted earlier, *urban and total population growth rates* generally move together, producing the *positively correlated relationship* we see here. The exception, also mentioned earlier, is China: there is a section where these two indicators are negatively correlated (i.e. total growth increased, but urban growth decreased).
Generally, it seems that there is a *negative correlation between life expectancy and total population growth* (though there is admittedly a lot of fluctuation). This could be because as people are able to live for longer periods of time, there’s less room for new people to be born or come in – so populations grow slower as existing residents age.
There seems to be a *positive correlation between life expectancy and population density.* As people live longer, there are more existing residents in a country at a given time. As babies continue to be born and immigrants continue to move in – but fewer people die out due to old age – the amount of people in a given area increases, i.e. population density.


## Session info

```{r}
devtools::session_info()
```


