---
title: "HW 4, pt. 2: Exploring World Bank Data"
author: "Julia Du"
output: github_document
---

## Load necessary libraries

```{r}
library(tidyverse)
library(ggthemes)
library(ggridges)
library(viridis)
theme_set(theme_minimal())

#don't do iteration w/in function, dummy!!
mean_mtcars <- vector("numeric", length = ncol(mtcars))
for(i in seq_along(mtcars)) {
  output[[i]] <- mean(mtcars[[i]])
}

max_diamonds <- vector("numeric", length = ncol(diamonds))
for(i in seq_along(diamonds)) {
  output[[i]] <- max(diamonds[[i]])
}
```

## Write a function to import the data files

  I'm interested in the following four variables: 
    
    1. Population growth (annual %) = SP.POP.GROW
    2. Urban population growth (annual %) = SP.URB.GROW
    3. Population density (people per sq. km of land area) = EN.POP.DNST
    4. "Life expectancy at birth, total (years)" = SP.DYN.LE00.IN

```{r}
import_tidydata <- function(file_name) {
 path = "./data_world_bank/"  #why this & not "~/data_world_bank/"
 read_csv(paste0(path, file_name), skip = 4) %>%
   rename(country_name = `Country Name`, country_code = `Country Code`, indicator_name = `Indicator Name`, indicator_code = `Indicator Code`) %>%
   filter(indicator_code %in% c("SP.POP.GROW", "SP.URB.GROW", "EN.POP.DNST", "SP.DYN.LE00.IN")) %>%
   #selecting chosen 4 variables: population growth, urban population growth, population density, & life expectancy at birth. doing it here & not in final select of function so that function runs faster
   select(-c(country_code, indicator_name)) %>%
   pivot_longer(cols = `1960`:`2019`, names_to = "year", values_to = "values") %>%
   pivot_wider(names_from = indicator_code, values_from = values) %>%
   select(-X65)
 }

#This is what I used to check my work while creating the function.
sampledfname <- "API_ABW_DS2_en_csv_v2_1226885.csv"
sampledf <- import_tidydata(sampledfname)
```

## Import the data

```{r}
masterdata <- list.files(path = "./data_world_bank/", pattern = "\\.csv$") %>% #could also write "*.csv"
map(import_tidydata) %>%
  reduce(rbind)
```

## Explore the data
I am focusing my analysis on Japan, China, & S. Korea.
```{r}
focus_data <- masterdata %>%
  filter(country_name %in% c("Japan", "China", "Korea, Rep.")) %>%
  mutate(country_name = if_else(country_name == "Korea, Rep.", "South Korea", country_name)) %>%
  rename(totalpopgrow = `SP.POP.GROW`, urbpopgrow = `SP.URB.GROW`, popdensity = `EN.POP.DNST`, life_expectancy = `SP.DYN.LE00.IN`)

### mostly looking at population variables

#comparing population growth across countries
ggplot(data = focus_data) +
  geom_point(mapping = aes(x = year, y = totalpopgrow/100, color = country_name, group = 1)) +
    labs(title = "Total Population Growth, 1960-2020", x = "Year", y = "Percent of total annual population growth", caption = "Source: The World Bank", color = "Country") +
  scale_x_discrete(breaks = seq(1960, 2020, by = 10)) +
  scale_y_continuous(labels = scales::percent)

#same as above, but with a line!!!
ggplot(data = focus_data) +
  geom_line(mapping = aes(x = year, y = totalpopgrow/100, color = country_name, group = country_name)) + 
    labs(title = "Total Population Growth, 1960-2020", x = "Year", y = "Percent of total annual population growth", caption = "Source: The World Bank", color = "Country") +
  scale_x_discrete(breaks = seq(1960, 2020, by = 10)) +
  scale_y_continuous(labels = scales::percent)

#compare urban vs total pop growth across countries
ggplot(data = focus_data, mapping = aes(x = year)) +
  geom_line(mapping = aes(y = totalpopgrow/100, color = "Total population growth"), group = 1) +
  geom_line(mapping = aes(y = urbpopgrow/100, color = "Urban population growth"), group = 1) +
  facet_grid(country_name ~.) +
    labs(title = "Urban & Total Population Growth, 1960-2020", x = "Year", y = "Percent of annual population growth", caption = "Source: The World Bank", color = "Legend") +
  scale_x_discrete(breaks = seq(1960, 2020, by = 10)) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "bottom")
```

```{r}
## looking @ life expectancy

#boxplot of life expectancy across country
ggplot(data = focus_data, mapping = aes(x = country_name, y = life_expectancy, fill = country_name)) +
  geom_violin() +
  geom_boxplot(width=0.1, color="grey", alpha=0.2) +
    labs(title = "Life expectancy in selected East Asian countries", x = "", y = "Life expectancy at birth (yrs)", caption = "Source: The World Bank", fill = "Country") +
  scale_y_continuous(limits = c(40, 90)) +
  theme(legend.position = "none")
#Japan has the highest life expectancy. Korea and China seem to have similar median life expectancies, but China is much more left skewed. 

ggplot(data = focus_data, mapping = aes(x = life_expectancy, y = country_name, fill = country_name)) +
  geom_density_ridges(alpha = 0.6) +
  scale_fill_viridis(discrete = TRUE) +
  scale_color_viridis(discrete = TRUE) +
    labs(title = "Life expectancy in selected East Asian countries", x = "Life expectancy at birth (yrs)", y = "", caption = "Source: The World Bank", fill = "Country") +
#  scale_x_continuous(limits = c(40, 100)) +
  theme(legend.position = "none")

#create scatter plot comparing life expectancy to population density
ggplot(focus_data, mapping = aes(x = life_expectancy, y = popdensity, color = country_name)) +
  geom_line() 


```

## Session info

```{r}
devtools::session_info()
```


